---
title: "Data Manipulation with the sparklyr Package"
output: html_document
---

# Create Spark Context

The `sparklyr` package has a handy function for creating a Spark context. This differs from the method that is used by the `SparkR` package.

```{r spark_context}
if ("SparkR" %in% loadedNamespaces()) detach("package:SparkR", unload = TRUE)

library(sparklyr)
sc <- spark_connect(master = "yarn-client")

```

# View Data Directory

```{r download_data}

wasb_taxi <- "/user/RevoShare/remoteuser/nyctaxi/data/"
hdfs_ls <- paste0("hadoop fs -ls ", wasb_taxi)
system(hdfs_ls)
```



# Import Data

To import data from csv files, we can use the `spark_read_csv` function, which is basically a wrapper for the `read.df` function using the __databricks.spark.csv__ package.

```{r import_csv}

taxi <- spark_read_csv(sc,
                       path = wasb_taxi,
                       "taxisample",
                       header = TRUE)


```


## Exploratory Data Analysis

```{r counts_year}
library(dplyr)

taxi %>% count

year_counts <- taxi %>% 
  mutate(year = year(tpep_pickup_datetime )) %>%
  group_by(year) %>%
  summarize(num_trips = n()) %>%
  collect()

monthly_counts <- taxi %>% 
  mutate(year = year(tpep_pickup_datetime ),
         month = month(tpep_pickup_datetime )) %>% 
  group_by(month, year) %>% 
  summarize(num_trips = n()) %>% 
  collect()

```

## Visualize Trip Durations

```{r}

library(ggplot2)

pickup_dropoff_tbl <- taxi %>%
  mutate(pickup_hour = hour(tpep_pickup_datetime)) %>%
  mutate(trip_time = unix_timestamp(tpep_dropoff_datetime) - unix_timestamp(tpep_pickup_datetime)) %>%
  group_by(pickup_hour) %>% 
  summarize(n = n(),
            trip_time_mean = mean(trip_time),
            trip_time_p10 = percentile(trip_time, 0.10),
            trip_time_p25 = percentile(trip_time, 0.25),
            trip_time_p50 = percentile(trip_time, 0.50),
            trip_time_p75 = percentile(trip_time, 0.75),
            trip_time_p90 = percentile(trip_time, 0.90))
# Collect results
pickup_dropoff <- collect(pickup_dropoff_tbl)
# Plot
ggplot(pickup_dropoff, aes(x = pickup_hour)) +
          geom_line(aes(y = trip_time_p50, alpha = "Median")) +
          geom_ribbon(aes(ymin = trip_time_p25, ymax = trip_time_p75, 
                          alpha = "25–75th percentile")) +
          geom_ribbon(aes(ymin = trip_time_p10, ymax = trip_time_p90, 
                          alpha = "10–90th percentile")) +
          scale_y_continuous("trip duration in minutes")
```

